---
title: "00_3 corpus compilation"
format: md
editor: visual
---

```{r}
library(tidyverse)
# library(tidytext)
```

# load data

## periodicals

Read Rds file with metadata & raw texts

```{r}
corpus_per <- readRDS("../../data/corpus1835/periodicals/per_corpus.Rds")

glimpse(corpus_per)
```

```{r}
# rename column
corpus_per <- corpus_per %>% 
  rename(text_raw = text) 
```

### attach lemmatized texts

```{r}
per_lemm <- read.csv("../../data/corpus1835/periodicals/per_corpus_lemm.csv")
glimpse(per_lemm)

corpus_per <- corpus_per %>% 
  left_join(per_lemm %>% select(text_id, text_lemm),
            by = "text_id")

rm(per_lemm)
```

Check if all texts are available (both outputs should be 0 rows)

```{r}
corpus_per %>% 
  filter(text_raw == "" | is.na(text_raw))

corpus_per %>% 
  filter(text_lemm == "" | is.na(text_lemm))
```

### attach acc texts

```{r}
# list files from accented dir
lf <- list.files(path = "../../data/corpus1835/periodicals/accented/",
                 full.names = T,
                 pattern = "\\.txt")

# read files and extract text_id for join
acc_corpus <- tibble(
  path = lf,
  text_acc = sapply(lf, read_file),
  text_id = str_extract(lf, "P_\\d+")
)

corpus_per <- corpus_per %>% 
  left_join(acc_corpus %>% select(text_id, text_acc), by = "text_id")

rm(acc_corpus, lf)
```

Check if all texts are available

```{r}
corpus_per %>% 
  filter(text_acc == "" | is.na(text_acc))
```

### n lines check

```{r}
# count number of lines in each text columns

raw_lines <- corpus_per %>% 
  select(text_id, text_raw) %>% 
  separate_rows(text_raw, sep = "\n") %>% 
  filter(text_raw != "") %>% 
  count(text_id) %>% 
  rename(nlines_raw = n)

head(raw_lines)

lem_lines <- corpus_per %>% 
  select(text_id, text_lemm) %>% 
  separate_rows(text_lemm, sep = "\n") %>% 
  filter(text_lemm != "") %>% 
  count(text_id) %>% 
  rename(nlines_lemm = n)

acc_lines <- corpus_per %>% 
  select(text_id, text_acc) %>% 
  separate_rows(text_acc, sep = "\n") %>% 
  filter(text_acc != "") %>% 
  count(text_id) %>% 
  rename(nlines_acc = n)

head(acc_lines)

# join results

nlines <- raw_lines %>% 
  left_join(lem_lines, by = "text_id") %>% 
  left_join(acc_lines, by = "text_id") 

errors_acc <- nlines %>% 
  filter(nlines_raw != nlines_acc)

errors_acc
```

rewrite errors, if needed

```{r, eval = F}
rewrite_errors <- corpus_per %>% 
  select(path, text_id, text_cln) %>% 
  filter(text_id %in% errors_acc$text_id) %>% 
  mutate(path_new = str_replace(path, "per_raw", "err_accented"), # create this folder before run!
         text_cln = str_replace(text_cln, "-", " - "),
         text_cln = str_remove_all(text_cln, "'"))

for (i in 1:nrow(rewrite_errors)) {
  write_file(rewrite_errors$text_cln[i], file = rewrite_errors$path_new[i])
}
```

### \<\> cleaning

Remove everything in \<\> as it's not parts of the texts

```{r}
corpus_per <- corpus_per %>% 
  mutate(text_cln = str_remove_all(text_raw, "<.*?>"),
         text_lemm = str_remove_all(text_lemm, "<.*?>"),
         text_acc = str_remove_all(text_acc, "<.*?>")) 

# corpus_per %>% 
#   filter(str_detect(text_acc, ">"))

# reorder columns

corpus_per <- corpus_per %>% 
  rename(text_subtitle = subtitle,
         text_pages = pages) %>% 
  select(text_id,
         A_ID,
         author_text, author_sign,
         # text data
         text_title, first_line, text_subtitle, 
         # text source data
         PER_ID, year, vol, num, text_pages,
         notes, path,
         # text versions 
         text_raw, text_cln, text_lemm, text_acc
         ) 
```

### save

```{r, eval = F}
# save all texts in a .csv

corpus_texts <- corpus_per %>% 
  select(text_id, text_raw, text_cln, text_lemm, text_acc)

write.csv(corpus_texts, "../../data/corpus1835/periodicals/per_texts_lemm_acc.csv")

# save all periodicals data in Rds
saveRDS(corpus_per, file = "../../data/corpus1835/periodicals/periodicals_corpus_lemm_acc.Rds")
```

Shorten metadata for merging with book data

```{r}
glimpse(corpus_per)
```

## books

Load Rds with metadata & raw texts

```{r}
corpus_books <- readRDS("../../data/corpus1835/collections/corpus_with_meta.Rds")
glimpse(corpus_books)
```

```{r}
# rename text to text_raw
corpus_books <- corpus_books %>% 
  rename(text_raw = text)
```

### attach lemmatized texts

```{r}
# read lemmatized
corpus_lemm <- read.csv("../../data/corpus1835/collections/books_corpus_lemm.csv")

corpus_books <- corpus_books %>% 
  left_join(corpus_lemm %>% select(text_id, text_lemm), by = "text_id")
```

Check

```{r}
corpus_books %>% 
  filter(text_raw == "" | is.na(text_raw))

corpus_books %>% 
  filter(text_lemm == "" | is.na(text_lemm))
```

```{r}
glimpse(corpus_per)
glimpse(corpus_books)
```

### attach acc texts

```{r}
# list files from accented dir
lf <- list.files(path = "../../data/corpus1835/collections/accented/",
                 full.names = T,
                 pattern = "\\.txt")

# read files and extract text_id for join
acc_corpus <- tibble(
  path = lf,
  text_acc = sapply(lf, read_file),
  text_id = str_extract(lf, "C_\\d+__\\d+")
)

corpus_books <- corpus_books %>% 
  left_join(acc_corpus %>% select(text_id, text_acc), by = "text_id")

rm(acc_corpus, lf)
```

```{r}
corpus_books %>% 
  filter(text_acc == "" | is.na(text_acc))
```

### n lines check

```{r}
raw_lines <- corpus_books %>% 
  select(text_id, text_raw) %>% 
  separate_rows(text_raw, sep = "\n") %>% 
  filter(text_raw != "") %>% 
  count(text_id) %>% 
  rename(nlines_raw = n)

head(raw_lines)

lem_lines <- corpus_books %>% 
  select(text_id, text_lemm) %>% 
  separate_rows(text_lemm, sep = "\n") %>% 
  filter(text_lemm != "") %>% 
  count(text_id) %>% 
  rename(nlines_lemm = n)

acc_lines <- corpus_books %>% 
  select(text_id, text_acc) %>% 
  separate_rows(text_acc, sep = "\n") %>% 
  filter(text_acc != "") %>% 
  count(text_id) %>% 
  rename(nlines_acc = n)

head(acc_lines)

# join results

nlines <- raw_lines %>% 
  left_join(lem_lines, by = "text_id") %>% 
  left_join(acc_lines, by = "text_id") 

errors_acc <- nlines %>% 
  filter(nlines_raw != nlines_acc)

errors_acc
```

### \<\> cleaning

```{r}
corpus_books <- corpus_books %>% 
  mutate(text_cln = str_remove_all(text_raw, "<.*?>"),
         text_lemm = str_remove_all(text_lemm, "<.*?>"),
         text_acc = str_remove_all(text_acc, "<.*?>")) 
```

### save

```{r, eval = F}
# save all texts in a .csv

corpus_texts <- corpus_books %>% 
  select(text_id, text_raw, text_cln, text_lemm, text_acc)

write.csv(corpus_texts, "../../data/corpus1835/collections/books_texts_lemm_acc.csv")

# save all periodicals data in Rds
saveRDS(corpus_books, file = "../../data/corpus1835/collections/books_corpus_lemm_acc.Rds")
```

# merge corpora

Take prepared books & periodicals corpora and merge them together (metadata shortened)

### books prep

```{r}
corpus_books <- readRDS("../../data/corpus1835/collections/books_corpus_lemm_acc.Rds")

glimpse(corpus_books)
```

Shorten the metadata

```{r}
corpus_books[is.na(corpus_books)] <- ""

corpus_books_short <- corpus_books %>% 
  # subcorpus tag (either per or col)
  mutate(corpus = "col") %>% 
  
  # paste text source
  mutate(source_text = paste0(author_book, " ", book_title, ". ", city, ": ", 
                         publisher, ", ", year, ". C. ", text_pages)) %>% 
  
  # select & reorder columns
  select(text_id, A_ID,
         author_text, author_sign,
         text_title, text_subtitle, first_line, 
         year, path_text, 
         source_text, COL_ID, corpus, 
         
         text_raw, text_cln, text_lemm, text_acc
         )

glimpse(corpus_books_short)
```

### per prep

Load data

```{r}
corpus_per <- readRDS("../../data/corpus1835/periodicals/periodicals_corpus_lemm_acc.Rds")

glimpse(corpus_per)
```

Shortening

```{r}
corpus_per[is.na(corpus_per)] <- ""

# create a new variable for merging
corpus_per_short <- corpus_per %>% 
  mutate(source_text = paste0(PER_ID, ". ", year),
         source_text = ifelse(vol != "", 
                              paste0(source_text, ". ", vol),
                              source_text),
         
         source_text = ifelse(num != "", 
                              paste0(source_text, ". №", num),
                              source_text),
         source_text = paste0(source_text, ". C. ", text_pages)) %>% 
  # add 'corpus' column & col_id (needed for books corpus sometimes)
  mutate(corpus = "per",
         COL_ID = "") %>% 
  # rename to join with corpus_books true paths
  rename(path_text = path) %>% 
  # select & reorder
  select(text_id, A_ID,
         author_text, author_sign,
         text_title, text_subtitle, first_line, 
         year, path_text, 
         source_text, COL_ID, corpus, 
         
         text_raw, text_cln, text_lemm, text_acc
         )

glimpse(corpus_per_short)
```

### merge

```{r}
corpus_1835 <- rbind(corpus_per_short, corpus_books_short)
```

### save

Save whole corpus as Rds

```{r, eval = F}
saveRDS(corpus_1835, file = "../../data/corpus1835/corpus1835_nometrics.Rds")
```

Save texts & metadata separately in csv

```{r, eval = F}
corpus1835_meta <- corpus_1835 %>% 
  select(-text_raw, -text_cln, -text_lemm, -text_acc)

write.csv(corpus1835_meta, file = "../../meta/texts_all_corpus1835.csv")

corpus1835_texts <- corpus_1835 %>% 
  select(text_id, text_raw, text_cln, text_lemm, text_acc)

write.csv(corpus1835_texts, file = "../../data/corpus1835/corpus1835_texts_lemm_acc.csv")
```
