---
title: "00_metrical_annotation"
format: html
editor: visual
---

## Load packages

```{r}
library(tidyverse)
library(tidytext)
```

```{r}
library(e1071)

source("classifier/calculate_regularity.R")
classifier <- readRDS("classifier/fitted_svm_radial.Rds")
```

## Load data

```{r}
load("../../data/corpus1835/corpus_1835_nometrics.Rda")
glimpse(corpus_1835)
```

Check if some texts are not accented

```{r}
corpus_1835 %>% 
  select(text_id, text_raw, text_lemm, text_acc) %>% 
  filter(is.na(text_acc)|text_acc == "")
```

```{r}
s <- corpus_1835 %>% 
  select(text_id, text_acc) %>% 
  rename(doc_id = text_id)

glimpse(s)
```

## Accents to binary notation

Accented texts to binary notation (Acce'nted -\> 010)

```{r}
binarised <- to_binary(s, document = "doc_id", text = "text_acc")
# 161,754 lines

glimpse(binarised)
```

## Meter labels

Count intervals freq for each text

```{r}
i <- extract_intervals(binarised,
                       no_anacrusis = F,
                       no_clausula = T)

head(i)

# nb this function wants 'doc_id' column
int_count <- rhythm_inequality(i,
                               raw_values = T,
                               drop_rare = T)

head(int_count)
```

Pivot for prediction

```{r}
dat <- int_count %>% 
  #ungroup() %>% 
  pivot_wider(names_from = no_stress,
              values_from = n, 
              values_fill = 0) %>% 
  select(-doc_id)

head(dat)
```

```{r}
p <- predict(classifier, newdata = dat[,-1])

head(p)
```

Percentages of meters (preliminary)

```{r}
tibble(doc_id = dat$doc_id,
       meter = p) %>% 
  count(meter) %>% 
  mutate(perc = round(n/4147*100, 1) )
```

## Feet annotation

```{r}
head(binarised)

feet_reg <- binarised %>% 
  # remove clausulas
  mutate(n_syl = str_replace_all(stress_pattern, "0+$", "")) %>% 
  mutate(n_syl = nchar(n_syl)) %>% 
  group_by(doc_id, n_syl) %>%
  count() %>%
  ungroup() %>%
  group_by(doc_id) %>%
  summarise(doc_id = doc_id,
            n_syl = n_syl, 
            n_lines = n,
            perc_feet = n/sum(n)) %>%
  filter(perc_feet > 0.6) %>% 
  slice_max(n = 1, perc_feet) %>% 
  ungroup() %>% 
  select(-n_lines)

head(feet_reg)
nrow(feet_reg) # 3,080 regular feet poems

feet_reg %>% 
  ggplot(aes(y = perc_feet)) + geom_boxplot()

feet_reg %>% 
  filter(is.na(n_syl))
```

Unite meters & syllable number

```{r}
glimpse(s)

s_m_syl <- tibble(doc_id = dat$doc_id, 
              meter = p) %>% # return data from the meter classifier
  left_join(s, by = "doc_id") %>% # attach accented texts
  left_join(feet_reg, by = "doc_id") # attach syllables count

glimpse(s_m_syl)

unique(s_m_syl$meter)
```

```{r}
# combine meters & feet
metrical_annotation <- s_m_syl %>% 
  
  # add number for feet division (n_syl / 2 for binary, n_syl/3 for ternary)
  mutate(binary_ternary = ifelse(meter %in% c("Iamb", "Trochee"), 2, 3)) %>% 
  
  # return removed clausulas for feet count
  mutate(n_syl = ifelse(meter %in% c("Trochee", "Amphibrach"), n_syl+1, n_syl),
         n_syl = ifelse(meter == "Dactyl", n_syl+2, n_syl),
         
         # divide for number of feets
         feet = n_syl/binary_ternary,
         
         # mark non-regular as "other", store regularity in a sep column
         feet = ifelse(is.na(feet), "other", feet),
         feet = ifelse(meter == "Other?", "?", feet)) %>% 
  mutate(feet_reg = ifelse(is.na(feet), "unreg", "reg"),
         
         # create 'formula' column
         formula = paste0(meter, "_", feet),) %>% 
  
  # attach vector with stress annotation collapsed in one char line
  
  left_join(
    binarised %>% 
      ungroup() %>% 
      group_by(doc_id) %>% 
      summarise(stress_pattern = paste0(stress_pattern, collapse = ", ")),
    by = "doc_id"
    ) %>% 
  
  # select resulting columns only
  select(doc_id, text_acc, meter, feet, formula, feet_reg, stress_pattern)

head(metrical_annotation)
```

### Add n_lines

```{r}
# count number of lines
t <- metrical_annotation 
t_lines <- t %>% 
  separate_rows(text_acc, sep = "\n") %>% 
  filter(text_acc != "") %>% 
  count(doc_id) %>% 
  rename(n_lines = n) 

# some brief test to check distribution
t_lines %>% ggplot(aes(x = n_lines)) + geom_boxplot()
mean(t_lines$n_lines)
median(t_lines$n_lines)

# attach to metadata tibble
metrical_annotation <- metrical_annotation %>% 
  left_join(t_lines, by = "doc_id")

glimpse(metrical_annotation)
```

## Test samples

Make samples for annotation check

```{r}

# t <- metrical_annotation %>% 
#   select(-stress_pattern) %>% 
#   filter(n_lines < 50) %>% 
#   sample_n(50) 

# write.csv(t, "annot_test_2.csv")
```

Some cleaning

```{r}
m <- metrical_annotation
```

```{r}
m %>% 
  count(formula, sort = T) %>% 
  mutate(perc = round(n/4149*100, 1) ) 

top_meters <- m %>% 
  count(formula, sort = T) %>% 
  mutate(perc = round(n/4149*100, 1) ) %>% 
  slice_max(n = 20, order_by = n) %>% pull(formula)

top_meters
```

```{r}
half_formulas <- m %>% 
  filter(str_detect(formula, "\\.5")) %>% 
  count(formula, sort = T) %>% pull(formula)

half_formulas

# t <- m %>% 
#   filter(formula %in% half_formulas)
# 
# write.csv(t, "half_formulas_test.csv")
```

```{r}
# NB there are also trimeters with 3.666 etc., should be fixed later
half_transl <- tibble(
  formula = half_formulas,
  true_formula = c("Iamb_4", "Iamb_6", "Iamb_5", "Trochee_4", "Trochee_5", "Iamb_4", "Trochee_6")
)

m_cln <- m %>% 
  left_join(half_transl, by = "formula") %>% 
  mutate(formula = ifelse(!is.na(true_formula), true_formula, formula)) %>% 
  select(-true_formula)
```

```{r}
glimpse(m_cln)

metrical_annotation <- m_cln

metrical_annotation %>% 
  count(formula, sort = T) %>% 
  mutate(perc = round((n/4149)*100, 1))
```

## Save

```{r}
write.csv(metrical_annotation,
          "../../data/corpus1835/corpus1835_metrical_annotation.csv")
```

### Attach to main corpus

```{r}
glimpse(corpus_1835)
glimpse(metrical_annotation)

corpus_1835_metrics <- corpus_1835 %>% 
  left_join(metrical_annotation %>% 
              select(-text_acc, -feet_reg) %>% 
              rename(text_id = doc_id),
            by = "text_id") 
```

```{r}
glimpse(corpus_1835_metrics)

corpus_1835 <- corpus_1835_metrics
save(corpus_1835, file = "../../data/corpus1835/corpus_1835.Rda")
```
