---
title: "04_3_networks"
format: html
editor: visual
---

# 4.2. Networks building

## load pckg

```{r}
library(tidyverse)

# networks
library(igraph)
library(ggraph)
library(ggrepel)

# basic colouring and layout options
library(wesanderson)
library(patchwork)
theme_set(theme_minimal())
```

## load data

```{r}
# load model output
load("data/lda75_output.Rda")

# load code for edges calculation
source("04_fn_calculate_edges.R")

# function to extract topic lables:
get_topic_labels = function(x, n_labels = 10) {
  
  #takes beta-matrix as an input
  top_terms = x %>%
    group_by(topic) %>%
    top_n(20, beta) %>%
    ungroup() %>%
    arrange(topic, -beta)
  
  topic_labels = top_terms %>%
    group_by(topic) %>%
    top_n(n_labels, beta) %>%
    summarise(label=paste(term, collapse=" ")) %>%
    mutate(label = paste(topic, label, sep="_"))
  
  return(topic_labels)
}

# prepare gamma data
gamma_rnc <- gamma %>% 
    # select ONLY RNC connections & topics
    filter(corpus == "N") %>% 
    # add index for create_edges function
    # mutate(index = paste0(corpus, "_", id)) %>% 
    # fix typo
    rename(author = first_line)

gamma <- gamma %>% 
  rename(author = first_line)

glimpse(gamma)
glimpse(gamma_rnc)

```

### extract topic labels

```{r}
topic_labels <- get_topic_labels(beta, n_labels = 5)
head(topic_labels)

# write.csv(topic_labels, file = "data/topic_labels.csv")
```

# RNC

## Create edges

### compute edges

```{r}
edges_raw <- compute_edges(gamma_rnc,
                           gamma_thresh = 0.05,
                           time_slice = 5)

head(edges_raw)
```

### calc top meters

```{r}
# select top meters
meter_counts <- gamma %>% 
  mutate(year_span = floor(as.numeric(year)/5)*5) %>% 
  rename("meter" = "formula") %>% 
  filter(!str_detect(meter, "other")) %>% 
  group_by(year_span, meter) %>% 
  count(sort = T) %>% 
  ungroup() %>% 
  filter(n > 10 & year_span != 1850)

head(meter_counts)
```

```{r}
# group meter counts in a variable
top_meters <- meter_counts %>% 
  group_by(meter) %>%
  summarise(total = sum(n)) %>% 
  arrange(desc(total))

head(top_meters, 15)
```

### create links

```{r}
# create links
links <- edges_raw %>% 
  mutate(meter = formula) %>%
  # filter out only 8 most interesting meters
  filter(meter %in% top_meters$meter[1:8]) %>% 
  # group and count edges
  group_by(slice, meter, corpus) %>% 
  count(slice, meter, edge_id, source, target, sort = T) %>% 
  # filter non-frequent edges
  filter(n > 2) %>% 
  ungroup() 

# quick check
head(links)
     
```

### create edgelist

```{r}
edgelist <- links %>% 
  select(source, target, n, meter, slice, corpus) %>% 
  mutate(width = n/10) %>% 
  filter(slice != 1850) %>% 
  rename(N_connections = n) %>% 
  mutate(facet_title = slice)

nodelist <- tibble(source = unique(c(links$target, links$source))) %>% 
  mutate(idn = as.numeric(str_replace(source, "^([0-9].*?)_.*", "\\1"))) 

head(edgelist)
head(nodelist) 
nrow(nodelist)
```

## create network

```{r}
net <- graph_from_data_frame(d = edgelist, vertices = nodelist, directed = F)
net
```

### viz fn

```{r}
# function for network visualisation

# taken as input net, meter as a string & colour palette value

network <- function(network, meter_value, palette_v) { 
  ggraph(network, layout = "stress") +
    geom_edge_fan(aes(color = meter,
                      filter = meter %in% c(meter_value)
                      #,
                      #width = N_connections
                     ),
                      alpha = 0.5, width = 1) +
    geom_node_point() +
    #geom_node_text(aes(label=idn), 
    #               hjust=0.1, 
    #               vjust=-0.4, 
    #               size=3, 
    #               color="grey50") +
    theme_void() + 
    theme(strip.text = element_text(size = 12)) + 
    scale_edge_color_manual(values = palette_v, aesthetics = "edge_colour") #+ 
    #theme(legend.position = "None")
}
```

```{r}
unique(edgelist$meter)
```

### all RNC viz (check)

```{r}
all_nkrja <- network(net, c("iamb_6", "iamb_free", "iamb_4", "trochee_4"), 
                     c(wes_palette("Darjeeling1")[2], 
                       wes_palette("Moonrise3")[2],
                       wes_palette("Darjeeling1")[3], 
                       wes_palette("Darjeeling1")[1])) 
                

all_nkrja + facet_wrap(~slice, 
                        scales="free_x", 
                        drop=T,
                        ncol = 5) + 
    labs(edge_colour = "Meter") + 
    theme(strip.text = element_text(size = 16),
         legend.text = element_text(size = 14), 
         legend.title = element_text(size = 16),
         panel.spacing = unit(1.4, "lines"),
         legend.position = "bottom")
        
```

### iamb-4

```{r}
iamb_4 <- network(net, c("iamb_4"), wes_palette("Darjeeling1")[2])
    
iamb_4 + facet_wrap(~slice, 
                 scales="free_x", 
                 drop=T,
                 ncol = 5) + 
    theme(legend.position = "None",
         strip.text = element_text(size = 16),
         legend.text = element_text(size = 14), 
         legend.title = element_text(size = 16))
```

### iamb-6

```{r}
iamb_6 <- network(net, c("iamb_6"), wes_palette("Darjeeling1")[3])
iamb_6 + facet_wrap(~slice, 
                 scales="free_x", 
                 drop=T,
                 ncol = 5) + 
    theme(legend.position = "None",
         strip.text = element_text(size = 16),
         legend.text = element_text(size = 14), 
         legend.title = element_text(size = 16))
```

### trochee-4

```{r}
trochee_4 <- network(net, c("trochee_4"), wes_palette("Darjeeling1")[1])
trochee_4 + facet_wrap(~slice, 
                 scales="free_x", 
                 drop=T,
                 ncol = 5) + 
    theme(legend.position = "None",
         strip.text = element_text(size = 16),
         legend.text = element_text(size = 14), 
         legend.title = element_text(size = 16))
```

## Network stats

### density

Calculate network density taking separately each time slice and each meter. The density of a graph is the ratio of the number of edges and the number of possible edges.

```{r}
density_vector <- function(formula) {
    densities <- vector()
    e1 <- NULL
    for (i in 1:length(unique(edgelist$slice))) {
        e1 <- edgelist %>% 
            filter(meter == formula & slice == unique(edgelist$slice)[i] & corpus == "N")

            net_test <- graph_from_data_frame(d = e1, vertices = nodelist, directed = F)

            densities[i] <- edge_density(net_test, loops = F)
    }
    densities
}
```

```{r}
length(unique(edgelist$slice))
```

```{r}
density_table <- tibble(meter = c(rep("iamb_4", 15), rep("iamb_6", 15), 
                 rep("iamb_free", 15), rep("trochee_4", 15)),
      slice = rep(unique(edgelist$slice), 4),
      density = round(c(density_vector("iamb_4"),
                       density_vector("iamb_6"),
                       density_vector("iamb_free"), 
                       density_vector("trochee_4")), 3)) %>% 
    arrange(desc(-slice)) 

density_table %>% 
    # pivot for easier reading
    pivot_wider(names_from = slice, values_from = density)
```

```{r}
summary(density_table$density)

density_table %>% 
    ggplot(aes(x = meter, y = density, fill = meter)) + geom_boxplot() + 
    scale_fill_manual(values = c(wes_palette("Darjeeling1")[2], 
                                wes_palette("Darjeeling1")[4],
                                wes_palette("Darjeeling1")[3],
                                wes_palette("Darjeeling1")[1]))
```

# RNC vs marginal corpus

```{r}
glimpse(gamma)
```

```{r}
t <- gamma %>% 
  filter(index == "M_P_1733" & gamma > 0.04) %>% 
  left_join(topic_labels, by = "topic")

t

gamma %>% 
    filter(index == "M_P_1733") %>% 
    ungroup() %>% 
    ggplot(aes(x = topic, y = gamma)) + 
        geom_line() + 
        
        # threshold line
        geom_hline(yintercept = 0.044, 
                   linetype = 2,
                   linewidth = 1.5,
                   colour = wes_palette("Darjeeling1")[2]) + 
        
        geom_point(data = gamma %>% filter(index == "M_P_1733" & gamma > 0.04), 
                   size = 2) + 
  
        geom_label_repel(data = t %>% filter(topic == "73"),
                   aes(label = label ),
                   nudge_y = 0.055,
                   nudge_x = -0.5
                  ) + 
        geom_label_repel(data = t %>% filter(topic == "56"),
                   aes(label = label ),
                   nudge_y = 0.025,
                   nudge_x = -5
                  ) + 
        geom_label_repel(data = t %>% filter(!topic %in% c("56", "73")),
                   aes(label = label ),
                   nudge_y = 0.01,
                   nudge_x = 0.5
                  ) + 
        
        labs(x = "Topic", y = "Topic probability") + 
        expand_limits(x = 80, y = 0.15) + 
        scale_x_continuous(breaks = seq(10, 70, 10)) +
        scale_y_continuous(breaks = seq(0, 0.15, 0.025)) + 
        theme(axis.text = element_text(size = 14),
             axis.title = element_text(size = 16))

#ggsave("plots/demo.png", plot = last_plot(), dpi = 300,
#       bg = "white", width = 8, height = 6)
```

```{r}
topic_probs <- gamma %>%
  group_by(topic) %>%
  summarise(avg_gamma = mean(gamma)) %>% 
  ungroup() %>% 
  arrange(desc(avg_gamma))

print("Most probable topics (avg)")

head(topic_probs, 10)

topic_probs %>% 
  ggplot(aes(reorder(topic,-avg_gamma), avg_gamma)) + geom_col()

gamma %>% 
  mutate(year_span = floor(as.numeric(year)/5)*5) %>% 
  group_by(year_span, topic) %>% 
  summarise(gamma_avg = mean(gamma)) %>% 
  top_n(10) %>% 
  ggplot(aes(x = year_span, y = gamma_avg, fill = as.factor(topic))) + geom_col()
```

### sampling

Check distribution in number of texts

```{r}
gamma %>% 
  select(author, year, corpus, index) %>% 
  distinct() %>% 
  count(author, corpus, sort = T) %>% 
  filter(author != "" & corpus == "M")

gamma %>% 
  select(author, year, corpus, index) %>% 
  distinct() %>% 
  count(author, corpus, sort = T) %>% 
  filter(author != "" & corpus == "M") %>% summary

gamma %>% 
  select(author, year, corpus, index) %>% 
  distinct() %>% 
  count(author, corpus, sort = T) %>% 
  filter(author != "" & corpus == "N")

gamma %>% 
  select(author, year, corpus, index) %>% 
  distinct() %>% 
  count(author, corpus, sort = T) %>% 
  filter(author != "" & corpus == "N") %>% summary
```

Select top authors - rnc

```{r}
rnc_authors <- gamma %>% 
    filter(corpus == "N") %>% 
    mutate(decade = floor(as.numeric(year)/5)*5) %>% 
    filter(decade %in% c(1830, 1835)) %>% 
    group_by(author) %>% 
    summarise(n_gamma = sum(n()),
             n_texts = n_gamma/75) %>% 
    arrange(desc(n_texts))

rnc_authors %>% 
    ungroup() %>%
    summarise(total_texts = sum(n_texts))

rnc_authors <- rnc_authors %>% 
    mutate(percentage = n_texts/2519*100)

head(rnc_authors, 20)
summary(rnc_authors$n_texts)
```

```{r}
# filter overpresented authors
top_authors <- rnc_authors %>% 
    filter(n_texts > 30) %>% select(author) # 19 authors

top_authors

# randomly select 30 texts by each author
sampled_id <- gamma %>% 
    filter(corpus == "N") %>% 
    # select slice
    filter(year > 1829 & year < 1840) %>%
    select(index, author) %>% 
    distinct() %>% 
    filter(author %in% top_authors$author) %>% 
    group_by(author) %>% 
    sample_n(30) %>% 
    ungroup() %>%
    select(-author)

# retrieve gamma values for samples for these authors
gamma_top <- inner_join(sampled_id, gamma, by = "index", multiple = "all")
```

Select top authors - corpus_1830

```{r}
m_authors <- gamma %>% 
    filter(corpus == "M" & author != "") %>% 
    mutate(decade = floor(as.numeric(year)/5)*5) %>% 
    filter(decade %in% c(1830, 1835)) %>% 
    group_by(author) %>% 
    summarise(n_gamma = sum(n()),
             n_texts = n_gamma/75) %>% 
    arrange(desc(n_texts))

m_authors %>% 
    ungroup() %>%
    summarise(total_texts = sum(n_texts))

m_authors <- m_authors %>% 
    mutate(percentage = n_texts/3867*100)

head(m_authors, 20)
summary(m_authors$n_texts)

m_top_authors <- m_authors %>% 
  filter(n_texts > 30) %>% select(author) # 30 authors

sampled_id_m <- gamma %>% 
    filter(corpus == "M") %>% 
    # select slice
    filter(year > 1829 & year < 1840) %>%
    select(index, author) %>% 
    distinct() %>% 
    filter(author %in% m_top_authors$author) %>% 
    group_by(author) %>% 
    sample_n(30) %>% 
    ungroup() %>%
    select(-author)

m_gamma_top <- inner_join(sampled_id_m, gamma, by = "index", multiple = "all")
```

Merging

```{r}
# select texts from RNC by non-prevalent authors
gamma_nontop <- gamma %>% 
    filter(corpus == "N" & !author %in% top_authors$author) %>% 
    filter(year > 1829 & year < 1840)

# retrieve all the gamma values for texts from periodicals
m_gamma_nontop <- gamma %>%
    filter(corpus == "M" & !author %in% m_top_authors$author) %>%
    filter(year > 1829 & year < 1840)

# merge the three together
gamma1830 <- rbind(gamma_top, gamma_nontop, 
                   m_gamma_top, m_gamma_nontop)

head(gamma1830)

print("Number of texts in each corpus after sampling:")
gamma1830 %>% 
    select(index, corpus) %>% 
    distinct() %>% 
    group_by(corpus) %>% 
    count()
```

```{r}
gamma %>% 
  filter(year > 1829 & year < 1840) %>% 
  select(index, corpus) %>% 
  distinct() %>% 
  group_by(corpus) %>% 
  count()
```

```{r}
save(gamma1830, file = "data/gamma_1830_sampled.Rda")
```

### corpus stats

```{r}
glimpse(gamma1830)
```

```{r}
print("Total number of unique poems in marginals corpus (after sampling):")
gamma1830 %>% 
    filter(corpus == "M") %>%  
    select(index) %>% 
    distinct() %>%
    nrow()
```

### meter counts

#### iamb_other == iamb_free

```{r}
gamma1830 <- gamma1830 %>% 
  mutate(formula = ifelse(formula == "iamb_other", "iamb_free", formula)) 
```

```{r}
# temporary make all iamb_other == iamb_free

counts_main <- gamma1830 %>% 
    select(index, corpus, formula) %>% 
    filter(formula %in% c("iamb_4", "iamb_6", "iamb_free", "trochee_4")) %>% 
    distinct() %>% 
    group_by(corpus, formula) %>% 
    count(sort = T) %>% 
    pivot_wider(names_from = formula, values_from = n)

rnc_others <- gamma1830 %>% 
    select(index, corpus, formula) %>% 
    distinct() %>% 
    filter(corpus == "N" & !formula %in% c("iamb_4", "iamb_6", "iamb_free", "trochee_4")) %>% 
    count() %>% 
    pull()

m_others <- gamma1830 %>% 
    select(index, corpus, formula) %>% 
    distinct() %>% 
    filter(corpus == "M" & !formula %in% c("iamb_4", "iamb_6", "iamb_free", "trochee_4")) %>% 
    count() %>% 
    pull()

counts <- cbind(counts_main, other = c(rnc_others, m_others))

counts_fin <- counts %>% 
    ungroup() %>% 
    rowwise(corpus) %>% mutate(total = sum(across(where(is.numeric))))

counts_fin

counts_fin[2,7][[1]]

counts_fin %>% 
    pivot_longer(!corpus, names_to = "formula", values_to = "n") %>% 
    group_by(corpus, formula) %>% 
    mutate(perc = ifelse(corpus == "N",
               round(n/counts_fin[2,7][[1]]*100,1),
               round(n/counts_fin[1,7][[1]]*100,1)))

```

### network

### topic labels & edges

```{r}
topic_labels <- get_topic_labels(beta, n_labels = 5)
head(topic_labels)

edges_1830 <- compute_edges(gamma1830,
                           gamma_thresh = 0.05,
                           time_slice = 5)

head(edges_1830)
```

```{r}
# select top meters
meter_counts_1830 <- gamma1830 %>% 
  rename("meter" = "formula") %>% 
  filter(!str_detect(meter, "other") & meter != "NA") %>% 
  group_by(meter, corpus) %>% 
  count(sort = T) %>% 
  mutate(n_texts = n/75) %>%
  ungroup() %>% 
  filter(n > 10)

top_meters_1830 <- meter_counts_1830 %>% 
  group_by(meter) %>%
  summarise(total = sum(n)) %>% 
  arrange(desc(total))

top_meters_1830$meter[1:8]
```

```{r}
links_1830 <- edges_1830 %>% 
  mutate(meter = formula) %>%
  # filter out only 8 most interesting meters
  filter(meter %in% top_meters_1830$meter[1:8]) %>% 
  # group and count edges
  group_by(meter, corpus) %>% 
  count(meter, edge_id, source, target, sort = T) %>% 
  # filter non-frequent edges
  filter(n > 2) %>% 
  ungroup() 

# quick check
head(links_1830)
```

```{r}
edges_rnc <- links_1830 %>% 
    mutate(meter_edge = paste0(meter, "__", edge_id)) %>% 
    filter(corpus == "N") %>% 
    pull(meter_edge)

edges_m <- links_1830 |> 
    mutate(meter_edge = paste0(meter, "__", edge_id)) %>% 
    filter(corpus == "M") %>% 
    pull(meter_edge)

head(edges_rnc)
head(edges_m)

edges_intersection <- intersect(edges_rnc, edges_m)
print("Intersection between built edges:")
head(edges_intersection)

length(edges_rnc)
length(edges_m)

length(edges_intersection)
```

### edgelist & nodelist

```{r}

edgelist <- links_1830 %>% 
    mutate(meter_edge = paste0(meter, "__", edge_id),
           # create a column indicating if an edge is present in both corpora
          group = ifelse(meter_edge %in% edges_intersection, "Both corpora", "Unique edge")) %>% 
    select(source, target, n, meter, corpus, group) %>% 
    mutate(width = n/10,
          corpus_fullname = ifelse(corpus == "M", "corpus_1830", "RNC"))

nodelist <- tibble(source = unique(c(links_1830$target, links_1830$source))) %>% 
  mutate(idn = as.numeric(str_replace(source, "^([0-9].*?)_.*", "\\1"))) 

head(edgelist)
head(nodelist) 
nrow(nodelist)
```

## Network viz

```{r}
net1830 <- graph_from_data_frame(d = edgelist, vertices = nodelist, directed = F)
net1830
```

### viz fn

```{r}
# function for network visualisation

# taken as input net, meter as a string & colour palette value

network <- function(network, meter_value, palette_v) {
  ggraph(network, layout = "kk") +
    # one layer for edges which are unique for each corpus (RNC or Periodicals)
    geom_edge_fan(aes(color = group,
                      filter = meter %in% c(meter_value) & group == "Unique edge", 
                      width = n), 
                      alpha = 0.4) +
    
    # second layer for intersected edges
    geom_edge_fan(aes(color = group,
                      filter = meter %in% c(meter_value) & group == "Both corpora", 
                      width = n), 
                      alpha = 1) +
    
    geom_node_point() +
    #geom_node_text(aes(label=idn), 
    #               hjust=0.1, 
    #               vjust=-0.4, 
    #               size=3, 
    #               color="grey50") +
    theme_void() + 
    theme(strip.text = element_text(size = 18)) +
    facet_wrap(~corpus_fullname, 
               scales="free_x", 
               drop=T,
               ncol = 2) + 
    scale_edge_color_manual(values = palette_v) + 
    #labs(title = meter_value) + 
    theme(legend.position = "None")
}
```

### iamb-4

```{r}
network(net1830, "iamb_4", c(wes_palette("Darjeeling1")[2], wes_palette("Royal1")[1]))
```

### iamb-6

```{r}
network(net1830, "iamb_6", c(wes_palette("Darjeeling1")[3], wes_palette("Royal1")[1]))
```

### trochee-4

```{r}
network(net1830, "trochee_4", c(wes_palette("Darjeeling1")[1], wes_palette("Royal1")[1]))
```

### comparison of connections

```{r}
links_n <- links_1830 %>% 
  filter(corpus == "N" & meter %in% c("iamb_4", "iamb_6", "iamb_free", "trochee_4")) %>% 
  mutate(connection = paste0(meter, "__", edge_id)) %>% 
  select(connection) 

links_m <- links_1830 %>% 
  filter(corpus == "M" & meter %in% c("iamb_4", "iamb_6", "iamb_free", "trochee_4")) %>% 
  mutate(connection = paste0(meter, "__", edge_id)) %>% 
  select(connection) 
  
t <- intersect(links_n$connection, links_m$connection)

head(t)
length(t)


```

```{r}
links_1830 %>% 
  filter(meter %in% c("iamb_4", "iamb_6", "iamb_free", "trochee_4")) %>% 
  mutate(connection = paste0(meter, "__", edge_id)) %>% 
  mutate(corpus_new = ifelse(connection %in% t, "Both corpora", corpus)) %>% 
  mutate(corpus_new = recode(corpus_new, 
                           "N" = "National corpus only", 
                           "M" = "Marginals only")) %>% 
  # first select unique connections and assign to both/P/N
  group_by(corpus_new, meter, edge_id) %>%
  count() %>% 
  select(-n) %>% 
  distinct() %>%
  # ungroup & count by large groups again
  ungroup() %>% 
  group_by(corpus_new, meter) %>% 
  count(sort = T) %>% 
  arrange(desc(meter)) %>% 
  ggplot(aes(x = meter,
           y = n,
           group = corpus_new,
           fill = corpus_new)) +
  geom_col(position = "dodge") +
  labs(x = "Meter",
       y = "", 
       fill = "Corpus",
       subtitle = "Number of unique connections found in RNC vs Periodicals") + 
  scale_fill_manual(values = c(wes_palette("Rushmore1")[3:5])) + 
  theme(axis.text = element_text(size = 11), 
        plot.subtitle = element_text(hjust = 0.5))
```

## No-sampling data

```{r}
topic_labels <- get_topic_labels(beta, n_labels = 5)
head(topic_labels)

glimpse(gamma)

gamma1830_ns <- gamma %>% 
  filter(year > 1829 & year < 1841)

gamma1830_ns %>% 
  select(index, corpus) %>% 
  distinct() %>% 
  count(corpus)

edges_1830_ns <- compute_edges(gamma1830_ns,
                           gamma_thresh = 0.05,
                           time_slice = 5)

head(edges_1830_ns)
```

```{r}
edges_1830_ns %>% 
  filter(index == "M_P_1733") 

gamma %>% 
  filter(index == "M_P_1733" & topic == 73) 


demo_edges <- edges_1830_ns %>% 
  ungroup() %>% 
  filter(index == "M_P_1733") %>% 
  select(source, target) 

demo_nodes <- tibble(source = unique(c(demo_edges$source, demo_edges$target)),
                     idn = str_extract(source, "^\\d+"))

demo_net <- graph_from_data_frame(demo_edges, demo_nodes, directed = F)

demo_graph <- ggraph(demo_net, "stress") + 
    geom_edge_fan(width = 1) + 
    geom_node_point(colour = wes_palette("Darjeeling1")[2], size = 14) + 
    geom_node_text(aes(label = idn), size = 8, colour = "white") + 
    scale_x_continuous(expand = c(0,0.7)) + 
    scale_y_continuous(expand = c(0, 0.7))

demo_graph
```

```{r}

top_meters <- c("iamb_4", "iamb_free", "trochee_4", "iamb_6")


links_1830_ns <- edges_1830_ns %>% 
  mutate(meter = formula) %>%
  # filter out only 8 most interesting meters
  filter(meter %in% top_meters) %>% 
  # group and count edges
  group_by(meter, corpus) %>% 
  count(meter, edge_id, source, target, sort = T) %>% 
  # filter non-frequent edges
  filter(n > 2) %>% 
  ungroup() 

# quick check
head(links_1830_ns)
```

```{r}
edges_rnc <- links_1830_ns %>% 
    mutate(meter_edge = paste0(meter, "__", edge_id)) %>% 
    filter(corpus == "N") %>% 
    pull(meter_edge)

edges_m <- links_1830_ns %>% 
    mutate(meter_edge = paste0(meter, "__", edge_id)) %>% 
    filter(corpus == "M") %>% 
    pull(meter_edge)

head(edges_rnc)
head(edges_m)

edges_intersection <- intersect(edges_rnc, edges_m)
print("Intersection between built edges:")
head(edges_intersection)

length(edges_rnc)
length(edges_m)
length(edges_intersection)

```

```{r}
edgelist_ns <- links_1830_ns %>% 
    mutate(meter_edge = paste0(meter, "__", edge_id),
           # create a column indicating if an edge is present in both corpora
          group = ifelse(meter_edge %in% edges_intersection, 
                         "Both corpora", 
                         "Unique edge")) %>% 
    select(source, target, n, meter, corpus, group) %>% 
    mutate(width = n/10,
          corpus_fullname = ifelse(corpus == "M", "Корпус-1835", "НКРЯ"))

nodelist_ns <- tibble(source = unique(c(links_1830_ns$target, links_1830_ns$source))) %>% 
  mutate(idn = as.numeric(str_replace(source, "^([0-9].*?)_.*", "\\1"))) 

head(edgelist_ns)
head(nodelist_ns) 
nrow(nodelist_ns)
```

```{r}
net1830_ns <- graph_from_data_frame(d = edgelist_ns, vertices = nodelist_ns, directed = F)
net1830_ns
```

```{r}
network(net1830_ns, "iamb_4", c(wes_palette("Darjeeling1")[2], wes_palette("Royal1")[1]))

ggsave("plots/Fig_4-2-4_iamb4.png", plot = last_plot(), bg = "white",
      dpi = 300, width = 8, height = 4)
```

```{r}
network(net1830_ns, "iamb_6", c(wes_palette("Darjeeling1")[3], wes_palette("Royal1")[1]))

ggsave("plots/Fig_4-2-4_iamb6.png", plot = last_plot(), bg = "white",
      dpi = 300, width = 8, height = 4)
```

```{r}
network(net1830_ns, "trochee_4", c(wes_palette("Darjeeling1")[1], wes_palette("Royal1")[1]))

ggsave("plots/Fig_4-2-4_trochee4.png", plot = last_plot(), bg = "white",
      dpi = 300, width = 8, height = 4)
```
