---
title: "03_2_projections"
format: md
editor: visual
---

# 3.1. Genres & words

This script analyses how words are distributed inside genres making UMAP projections of distances based on MFW.

```{r}
library(tidyverse)
library(tidytext)

library(umap)

library(MetBrewer)
theme_set(theme_minimal())
```

Load data

```{r}
corpus1835 <- readRDS("../../data/corpus1835/corpus_1835.Rds")

titles <- read.csv("poems_titles.csv") %>% select(-X)
```

Attach titles data to the corpus: for genres only

```{r}
genres <- unlist(str_split(c("песня
романс
сонет
элегия
басня
песнь
альбом
дума
баллада
отрывок
подражание
послание
молитва
фантазия
псалом
эпиграмма
мелодия
антологический
аполог
сказка
экспромт
надпись
эпилог"), pattern = "\n"))

multigenre <- titles %>% 
  select(text_id, lemma) %>% 
  filter(lemma %in% genres) %>% 
  group_by(text_id) %>% 
  summarise(genre = paste0(lemma, collapse = " | ")) %>% 
  filter(str_detect(genre, " \\| "))

m <- corpus1835 %>% 
  filter(text_id %in% multigenre$text_id) %>% 
  select(text_id, text_title, text_subtitle) %>% 
  left_join(multigenre %>% select(text_id, genre), by = "text_id")

# write.csv(m, "multigenre_texts.csv")
```

```{r}
# read the data with normalised genres
m <- read.delim("multigenre_texts.csv", sep = ';') %>% select(-X)

glimpse(m)

# attach normalised genres to other genre titles
ids_genres <- titles %>% 
  select(text_id, lemma) %>% 
  filter(lemma %in% genres) %>% 
  group_by(text_id) %>% 
  summarise(genre = paste0(lemma, collapse = " | ")) %>% 
  filter(!str_detect(genre, " \\| ")) %>% 
  rbind(m %>% 
          select(text_id, true_genre) %>% 
          rename(genre = true_genre)
          )
```

```{r}
# glimpse(corpus1835)

corpus_genres <- corpus1835 %>% 
  select(text_id, author_text, text_title, year, 
         text_cln, text_lemm, 
         meter, feet, n_lines) %>% 
  left_join(ids_genres, by = "text_id") %>% 
  mutate(genre = ifelse(genre == "", NA, genre))

table(corpus_genres$genre)
```

## texts with genre titles

```{r}
corpus_genres %>% 
  filter(!is.na(genre)) %>% 
  select(-text_lemm) %>% 
  separate_rows(text_cln, sep = "\n") %>% 
  filter(text_cln != "" & !str_detect(text_cln, "^\\W+$")) %>% 
  count(genre, text_id) %>% 
  group_by(genre) %>% 
  summarise(mean_lines = round(mean(n),2),
            median_lines = round(median(n),2)) %>% 
  ungroup() %>% 
  pivot_longer(!genre, names_to = "metric", values_to = "n_lines") %>% 
  ggplot(aes(x = reorder_within(genre, n_lines, genre), y = n_lines, fill = metric)) + 
  geom_col(position = "dodge") + 
  coord_flip() + 
  scale_x_reordered() + 
  scale_fill_manual(values = c(met.brewer("Veronese")[3],
                               met.brewer("Veronese")[6])) + 
  labs(x = "",
       y = "Количество строк")
```

```{r}
corpus_genres %>% 
  filter(!is.na(genre)) %>% 
  select(-text_lemm) %>% 
  separate_rows(text_cln, sep = "\n") %>% 
  filter(text_cln != "" & !str_detect(text_cln, "^\\W+$")) %>% 
  mutate(corpus = str_extract(text_id, "^\\w")) %>% 
  count(corpus, genre, text_id) %>% 
  group_by(genre, corpus) %>% 
  summarise(median_lines = round(median(n),2)) %>% 
  ungroup() %>% 
  ggplot(aes(x = reorder_within(genre, median_lines, genre), 
             y = median_lines, 
             fill = corpus)) + 
  geom_col(position = "dodge", width = 0.7) + 
  coord_flip() + 
  scale_x_reordered() + 
  scale_fill_manual(values = c(met.brewer("Veronese")[3],
                               met.brewer("Veronese")[6])) + 
  labs(x = "",
       y = "Количество строк")
```

## Sampling

Take random samples of 25 lines from all texts longer than 25, filter out texts shorter than 8 lines.

```{r}
source("03_fn_sampling.R")
```

```{r}
# filter out short poems (<7 lines)
corpus_genres <- corpus_genres %>% 
  filter(n_lines > 7)

# -260 poems
```

Sample poems which are too long

```{r}
genres_sampled <- sample_long(corpus_genres %>%  
                                #select(-text_lemm) %>% 
                                rename(#text_lemm = text_cln,
                                  id = text_id),
                        starting_n = 1,
                        sample_size = 25,
                        over9000 = 5)
```

Comparison on the number of genre-labelled poems before and after sampling

```{r}
corpus_genres %>% 
  filter(!is.na(genre)) %>% 
  nrow()

genres_sampled %>% 
  filter(!is.na(genre)) %>% 
  nrow()
```

Leave only the texts with genre titles

```{r}
genres_only <- genres_sampled %>% 
  filter(!is.na(genre))

glimpse(genres_only)
```

```{r}
genres_only %>% 
  #select(-text_lemm) %>% 
  separate_rows(text_lemm, sep = "\n") %>% 
  filter(text_lemm != "" & !str_detect(text_lemm, "^\\W+$")) %>% 
  count(id, genre) %>%  
  ggplot(aes(x = genre, y = n)) + geom_boxplot()
```

## Projection

```{r}
genres_only %>% 
  unnest_tokens(input = text_lemm, output = word, token = "words") %>% 
  group_by(genre) %>% 
  count(word, sort = T) %>% 
  slice_max(n, n = 10) %>% 
  ungroup() %>% 
  pivot_wider(names_from = genre, values_from = n)
```

### ranks

Extract MFW from the corpus

```{r}
ranks <- genres_only %>% 
  unnest_tokens(input = text_lemm, output = word, token = "words") %>% 
  count(word, sort = T) %>% 
  head(1500)

head(ranks, 10)

tail(ranks, 10)

ranks %>% 
  mutate(rank = row_number()) %>% 
  sample_n(15) %>% 
  arrange(-desc(rank))
```

Count N of MFW in each poem

```{r}
counter <- genres_only %>% 
  unnest_tokens(input = text_lemm, output = word, token = "words") %>% 
  group_by(id) %>% 
  count(word) %>% 
  filter(word %in% ranks$word) %>% 
  ungroup()

counter %>% 
  sample_n(10)
```

Create a matrix & make a UMAP projection

```{r}
xxx <- counter %>% 
  pivot_wider(names_from = word, values_from = n, values_fill = 0)

mtrx <- xxx %>% 
  ungroup() %>% 
  select(-id) %>% 
  scale()

dim(mtrx)

u <- umap(mtrx)

dat <- tibble(x = u$layout[,1],
       y = u$layout[,2],
       id = xxx$id) %>% 
  left_join(genres_only, by = "id") 
  
table(genres_only$genre)

glimpse(dat)
```

```{r}
dat %>% 
  filter(meter != "Other?") %>% 
  ggplot(aes(x, y, color = meter, shape = meter)) + 
  geom_point(size = 5, alpha = 0.6)
```

```{r}
dat %>% 
  filter(meter != "Other?") %>% 
  ggplot(aes(x, y, color = genre)) + 
  geom_point(size = 4, alpha = 0.6) + 
  theme(legend.position = "None")
```

```{r}
dat %>% 
  filter(meter != "Other?") %>% 
  ggplot(aes(x, y, color = author_text)) + 
  geom_point(size = 4, alpha = 0.6) + 
  theme(legend.position = "None")
```

```{r}
dat %>% 
  filter(meter != "Other") %>% 
  
  filter(meter == "Iamb") %>% 
  
  filter(genre == "элегия") %>% 
  
  ggplot(aes(x, y, color = genre, shape = meter)) + 
  geom_point(size = 5, alpha = 0.7) + 
  geom_text(aes(label = author_text), size = 4, color = "black") + 
  scale_color_manual(values = c(met.brewer("Archambault")))
```

```{r}
dat %>% 
  mutate(first_line = str_extract(text_cln, "^(.*?)\n"),
         author_text = paste0(first_line)) %>% 
  filter(meter != "Other") %>% 
  
  filter(meter == "Iamb" | 
    meter == "Trochee") %>% 
  
  filter(genre == "песня") %>% 
  
  ggplot(aes(x, y, color = meter, shape = meter)) + 
  geom_point(size = 5, alpha = 0.7) + 
  geom_text(aes(label = author_text), size = 4) + 
  scale_color_manual(values = c(met.brewer("Archambault")))
```

```{r}
dat %>% 
  #mutate(first_line = str_extract(text_cln, "^(.*?)\n"),
  #       author_text = paste0(first_line)) %>% 
  filter(meter != "Other") %>% 
  
  filter(meter == "Iamb" | 
    meter == "Trochee") %>% 
  
  filter(genre == "песня") %>% 
  
  ggplot(aes(x, y, color = meter, shape = meter)) + 
  geom_point(size = 5, alpha = 0.7) + 
  geom_text(aes(label = author_text), size = 4) + 
  scale_color_manual(values = c(met.brewer("Archambault")))
```

```{r}
dat %>% 
  filter(meter != "Other") %>% 
  
  filter(meter == "Trochee") %>% 
  
  filter(genre %in% c("песня",
                      "послание",
                      "басня",
                      "сонет",
                      "элегия",
                      "баллада",
                      "псалом")) %>%
  
  ggplot(aes(x, y, color = author_text)) + 
  geom_point(size = 5, alpha = 0.7) + 
  geom_text(aes(label = author_text), size = 4) + theme(legend.position = "None")
```

```{r}
dat %>% 
  filter(meter != "Other") %>% 
  
  filter(meter == "Trochee" | meter == "Iamb") %>% 
  
  filter(genre %in% c("песня",
                      "послание",
                      "басня",
                      "сонет",
                      "элегия",
                      #"баллада",
                      "псалом")) %>%
  
  ggplot(aes(x, y, color = genre, shape = meter)) + 
  geom_point(size = 5, alpha = 0.7) + 
  #geom_text(aes(label = author_text), size = 4) + theme(legend.position = "None")
  scale_color_manual(values = c(met.brewer("Archambault")))
```
