---
title: "03_2_classification"
format: md
editor: visual
---

## Classification

This notebook uses samples of different poetic genres from the Corpus-1835 and try to classify them using SVM. The aim is to test whether even basic MFW-based classification accuracy will be higher than chance, and if yes, what are the "influential" words behind this result.

### load data

```{r}
library(tidyverse)
library(tidytext)

library(e1071)
library(caret)

library(tidymodels)
library(textrecipes)
library(kernlab)

library(MetBrewer)
theme_set(theme_minimal())
```

```{r}
# main corpus
corpus1835 <- readRDS("../../data/corpus1835/corpus_1835.Rds")

# read genre titles
titles <- read.csv("poems_titles.csv") %>% select(-X)

# read the data with normalised genres
m <- read.delim("multigenre_texts.csv", sep = ';') %>% select(-X)

# list of genre words (longer)
genres <- unlist(str_split(c("песня
романс
сонет
элегия
басня
песнь
альбом
дума
баллада
отрывок
подражание
послание
молитва
фантазия
псалом
эпиграмма
мелодия
антологический
аполог
сказка
экспромт
надпись
эпилог"), pattern = "\n"))

# attach normalised genres to other genre titles
ids_genres <- titles %>% 
  select(text_id, lemma) %>% 
  filter(lemma %in% genres) %>% # select genres from the list
  group_by(text_id) %>% 
  summarise(genre = paste0(lemma, collapse = " | ")) %>% 
  filter(!str_detect(genre, " \\| ")) %>% # remove multigenre titles
  rbind(m %>% 
          select(text_id, true_genre) %>% 
          rename(genre = true_genre) # replace them with clean labels
          ) %>% 
  filter(genre != "")

rm(genres, m, titles)
```

```{r}
glimpse(ids_genres)
ids_genres %>% count(genre, sort = T)
```

Select only texts with genre titles & needed columns

```{r}
corpus_genres <- corpus1835 %>% 
  left_join(ids_genres, by = "text_id") %>% 
  #filter(!is.na(genre)) %>% 
  select(text_id, genre, meter, text_lemm) %>% 
  mutate(genre = ifelse(is.na(genre), "no_genre", genre))
```

## poem-level classification

### ranks

Count word ranks (lemmatised corpus) for all genres

NB here poems without genre titles are NOT filtered out, "others" are all possible other texts, both with and without genre titles.

```{r}
ranks <- corpus_genres %>% 
  unnest_tokens(input = text_lemm, output = word, token = "words") %>% 
  count(word, sort = T) %>% 
  mutate(rank = row_number()) %>% 
  head(300)

head(ranks, 10)
tail(ranks, 10)
```

### песня

Create genre lables

```{r}
corpus_lbld <- corpus_genres %>% 
  mutate(text_genre = ifelse(genre == "песня", genre, "другое")) 
# leave only genre title of the needed class, rename the rest as 'other'

table(corpus_lbld$text_genre) # number of available texts
```

Cound 300MFW freq in all texts in the labelled corpus

```{r}
freqs <- corpus_lbld %>% 
  select(text_id, text_genre, text_lemm) %>% 
  unnest_tokens(input = text_lemm, output = word, token = "words") %>% 
  filter(word %in% ranks$word) %>% 
  group_by(text_id, text_genre) %>% 
  count(word) %>% 
  ungroup() %>% 
  pivot_wider(names_from = word, values_from = n, values_fill = 0) 

dim(freqs)
freqs[1:10, 1:10]
```

Scaling

```{r}
freqs_scaled <- freqs %>% 
  select(-text_id, -text_genre) %>% 
  scale() %>% 
  as.tibble()


freqs_scaled[1:10, 1:10]

# glimpse(freqs_scaled)

freqs_scaled <- tibble(text_id = freqs$text_id,
       freqs_scaled,
       text_genre = freqs$text_genre) %>% 
  mutate(text_id = row_number())

# glimpse(freqs_scaled)
```

Use tidymodels for 10-fold cross validation & fitting

```{r}
tidy_corpus <- freqs_scaled %>% 
  group_by(text_genre) %>% 
  sample_n(242) %>%   # number of texts in the tested genre
  ungroup() %>% 
  select(-text_id)

## split the data into training and test sets
corpus_split <- initial_split(tidy_corpus, strata="text_genre", prop = 7.5/10)

corpus_split

training_set <- training(corpus_split)
test_set <- testing(corpus_split)

# colnames(training_set) == colnames(test_set)
```

Settings for the classifier

```{r}
zscores_recipe <- recipe(text_genre ~ ., data = training_set) # all columns = all 300MFW freqs are used for the training

# 10-fold cross validation setup
folds <- vfold_cv(training_set, strata = "text_genre", v = 10)

# model specifications
svm_specs <- svm_poly(cost=1,degree = 1) %>% # linear kernel
  set_mode("classification") %>%
  set_engine("kernlab")

# add recipe and model specs to the workflow 
svm_wf <- workflow() %>%
  add_recipe(zscores_recipe) %>%
  add_model(svm_specs)
```

Fitting SVM

```{r}
svm_res <- fit_resamples(
  svm_wf,
  folds,
  metrics = metric_set(accuracy),
  control = control_resamples(save_pred = TRUE)
)

```

```{r}
metrics = collect_metrics(svm_res)
metrics 
```

```{r}
svm_model <-svm(as.factor(text_genre)~.,  
                data = training_set, 
                method = "C-classification", 
                kernel = "linear", 
                cost = 1, 
                scale = T)

summary(svm_model) 
prediction <- predict(svm_model, test_set)
confusionMatrix(prediction, as.factor(test_set$text_genre)) # NB check if the same positive class used below in the plot

words_coefs <- t(svm_model$coefs) %*% svm_model$SV
```

```{r}
tibble(weight=words_coefs[1,], word=colnames(words_coefs)) %>% 
  mutate(genre = case_when(weight > 0 ~ "другое", 
                           weight < 0 ~ "песня")) %>%
  group_by(genre) %>% 
  mutate(abs=abs(weight)) %>%
  top_n(20,abs) %>% 
  ggplot(aes(reorder(word,abs), abs, fill=genre)) + geom_col() +
  coord_flip() + 
  facet_wrap(~genre,scales="free") +
  theme_minimal(base_size = 16) + 
  labs(x = "", 
       y = "",
       fill = "") + 
  #scale_fill_carto_d(palette = "Safe") + 
  theme(legend.position = "none") + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 14)) + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  theme(axis.line.x = element_line(color="black"), 
        axis.line.y = element_line(color="black"))
```

### романс

```{r}
corpus_lbld <- corpus_genres %>% 
  mutate(text_genre = ifelse(genre == "романс", genre, "другое")) 
# leave only genre title of the needed class, rename the rest as 'other'

table(corpus_lbld$text_genre) # number of available texts

freqs <- corpus_lbld %>% 
  select(text_id, text_genre, text_lemm) %>% 
  unnest_tokens(input = text_lemm, output = word, token = "words") %>% 
  filter(word %in% ranks$word) %>% 
  group_by(text_id, text_genre) %>% 
  count(word) %>% 
  ungroup() %>% 
  pivot_wider(names_from = word, values_from = n, values_fill = 0) 

# dim(freqs)
# freqs[1:10, 1:10]

freqs_scaled <- freqs %>% 
  select(-text_id, -text_genre) %>% 
  scale() %>% 
  as.tibble()


# freqs_scaled[1:10, 1:10]

# glimpse(freqs_scaled)

freqs_scaled <- tibble(text_id = freqs$text_id,
       freqs_scaled,
       text_genre = freqs$text_genre) %>% 
  mutate(text_id = row_number())


tidy_corpus <- freqs_scaled %>% 
  group_by(text_genre) %>% 
  sample_n(104) %>%   # number of texts in the tested genre
  ungroup() %>% 
  select(-text_id)

## split the data into training and test sets
corpus_split <- initial_split(tidy_corpus, strata="text_genre", prop = 7.5/10)

corpus_split

training_set <- training(corpus_split)
test_set <- testing(corpus_split)

# colnames(training_set) == colnames(test_set)

# fit the model
svm_res <- fit_resamples(
  svm_wf,
  folds,
  metrics = metric_set(accuracy),
  control = control_resamples(save_pred = TRUE)
)

metrics = collect_metrics(svm_res)
metrics 

# check the word coefficients with e1071
svm_model <-svm(as.factor(text_genre)~.,  
                data = training_set, 
                method = "C-classification", 
                kernel = "linear", 
                cost = 1, 
                scale = T)

summary(svm_model) 
prediction <- predict(svm_model, test_set)
confusionMatrix(prediction, as.factor(test_set$text_genre)) # NB check if the same positive class used below in the plot

words_coefs <- t(svm_model$coefs) %*% svm_model$SV

tibble(weight=words_coefs[1,], word=colnames(words_coefs)) %>% 
  mutate(genre = case_when(weight > 0 ~ "другое", 
                           weight < 0 ~ "романс")) %>%
  group_by(genre) %>% 
  mutate(abs=abs(weight)) %>%
  top_n(20,abs) %>% 
  ggplot(aes(reorder(word,abs), abs, fill=genre)) + geom_col() +
  coord_flip() + 
  facet_wrap(~genre,scales="free") +
  theme_minimal(base_size = 16) + 
  labs(x = "", 
       y = "",
       fill = "") + 
  #scale_fill_carto_d(palette = "Safe") + 
  theme(legend.position = "none") + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 14)) + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  theme(axis.line.x = element_line(color="black"), 
        axis.line.y = element_line(color="black"))
```

### элегия

```{r}
corpus_lbld <- corpus_genres %>% 
  mutate(text_genre = ifelse(genre == "элегия", genre, "другое")) 
# leave only genre title of the needed class, rename the rest as 'other'

table(corpus_lbld$text_genre) # number of available texts

freqs <- corpus_lbld %>% 
  select(text_id, text_genre, text_lemm) %>% 
  unnest_tokens(input = text_lemm, output = word, token = "words") %>% 
  filter(word %in% ranks$word) %>% 
  group_by(text_id, text_genre) %>% 
  count(word) %>% 
  ungroup() %>% 
  pivot_wider(names_from = word, values_from = n, values_fill = 0) 

# dim(freqs)
# freqs[1:10, 1:10]

freqs_scaled <- freqs %>% 
  select(-text_id, -text_genre) %>% 
  scale() %>% 
  as.tibble()


# freqs_scaled[1:10, 1:10]

# glimpse(freqs_scaled)

freqs_scaled <- tibble(text_id = freqs$text_id,
       freqs_scaled,
       text_genre = freqs$text_genre) %>% 
  mutate(text_id = row_number())


tidy_corpus <- freqs_scaled %>% 
  group_by(text_genre) %>% 
  sample_n(72) %>%   # number of texts in the tested genre
  ungroup() %>% 
  select(-text_id)

## split the data into training and test sets
corpus_split <- initial_split(tidy_corpus, strata="text_genre", prop = 7.5/10)

corpus_split

training_set <- training(corpus_split)
test_set <- testing(corpus_split)

# colnames(training_set) == colnames(test_set)

# fit the model
svm_res <- fit_resamples(
  svm_wf,
  folds,
  metrics = metric_set(accuracy),
  control = control_resamples(save_pred = TRUE)
)

metrics = collect_metrics(svm_res)
metrics 

# check the word coefficients with e1071
svm_model <-svm(as.factor(text_genre)~.,  
                data = training_set, 
                method = "C-classification", 
                kernel = "linear", 
                cost = 1, 
                scale = T)

summary(svm_model) 
prediction <- predict(svm_model, test_set)
confusionMatrix(prediction, as.factor(test_set$text_genre)) # NB check if the same positive class used below in the plot

words_coefs <- t(svm_model$coefs) %*% svm_model$SV

tibble(weight=words_coefs[1,], word=colnames(words_coefs)) %>% 
  mutate(genre = case_when(weight > 0 ~ "другое", 
                           weight < 0 ~ "элегия")) %>%
  group_by(genre) %>% 
  mutate(abs=abs(weight)) %>%
  top_n(20,abs) %>% 
  ggplot(aes(reorder(word,abs), abs, fill=genre)) + geom_col() +
  coord_flip() + 
  facet_wrap(~genre,scales="free") +
  theme_minimal(base_size = 16) + 
  labs(x = "", 
       y = "",
       fill = "") + 
  #scale_fill_carto_d(palette = "Safe") + 
  theme(legend.position = "none") + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 14)) + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  theme(axis.line.x = element_line(color="black"), 
        axis.line.y = element_line(color="black"))
```

### басня

```{r}
corpus_lbld <- corpus_genres %>% 
  mutate(text_genre = ifelse(genre == "басня", genre, "другое")) 
# leave only genre title of the needed class, rename the rest as 'other'

table(corpus_lbld$text_genre) # number of available texts

freqs <- corpus_lbld %>% 
  select(text_id, text_genre, text_lemm) %>% 
  unnest_tokens(input = text_lemm, output = word, token = "words") %>% 
  filter(word %in% ranks$word) %>% 
  group_by(text_id, text_genre) %>% 
  count(word) %>% 
  ungroup() %>% 
  pivot_wider(names_from = word, values_from = n, values_fill = 0) 

# dim(freqs)
# freqs[1:10, 1:10]

freqs_scaled <- freqs %>% 
  select(-text_id, -text_genre) %>% 
  scale() %>% 
  as.tibble()


# freqs_scaled[1:10, 1:10]

# glimpse(freqs_scaled)

freqs_scaled <- tibble(text_id = freqs$text_id,
       freqs_scaled,
       text_genre = freqs$text_genre) %>% 
  mutate(text_id = row_number())


tidy_corpus <- freqs_scaled %>% 
  group_by(text_genre) %>% 
  sample_n(69) %>%   # number of texts in the tested genre
  ungroup() %>% 
  select(-text_id)

## split the data into training and test sets
corpus_split <- initial_split(tidy_corpus, strata="text_genre", prop = 7.5/10)

corpus_split

training_set <- training(corpus_split)
test_set <- testing(corpus_split)

# colnames(training_set) == colnames(test_set)

# fit the model
svm_res <- fit_resamples(
  svm_wf,
  folds,
  metrics = metric_set(accuracy),
  control = control_resamples(save_pred = TRUE)
)

metrics = collect_metrics(svm_res)
metrics 

# check the word coefficients with e1071
svm_model <-svm(as.factor(text_genre)~.,  
                data = training_set, 
                method = "C-classification", 
                kernel = "linear", 
                cost = 1, 
                scale = T)

summary(svm_model) 
prediction <- predict(svm_model, test_set)
confusionMatrix(prediction, as.factor(test_set$text_genre)) # NB check if the same positive class used below in the plot

words_coefs <- t(svm_model$coefs) %*% svm_model$SV

tibble(weight=words_coefs[1,], word=colnames(words_coefs)) %>% 
  mutate(genre = case_when(weight > 0 ~ "басня", 
                           weight < 0 ~ "другое")) %>%
  group_by(genre) %>% 
  mutate(abs=abs(weight)) %>%
  top_n(20,abs) %>% 
  ggplot(aes(reorder(word,abs), abs, fill=genre)) + geom_col() +
  coord_flip() + 
  facet_wrap(~genre,scales="free") +
  theme_minimal(base_size = 16) + 
  labs(x = "", 
       y = "",
       fill = "") + 
  #scale_fill_carto_d(palette = "Safe") + 
  theme(legend.position = "none") + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 14)) + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  theme(axis.line.x = element_line(color="black"), 
        axis.line.y = element_line(color="black"))
```

## samples classification

Below aggregated samples are used for the classification, not individual texts. One sample is 20 lines, 50 samples taken from each genre (including 'no_genre' class).

```{r}
corpus_genres %>% count(genre, sort = T) %>% head()

genres_sampled <- corpus_genres %>%
  
  # filter only genres with > 1000 lines available
  filter(genre %in% c("песня", "романс", "элегия", "баллада", "послание", 
                      "песнь", "отрывок", "сонет", "басня", 
                      "дума", "no_genre")) %>% 
  
  separate_rows(text_lemm, sep = "\n") %>% 
  group_by(genre) %>% 
  sample_n(1000) %>% 
  mutate(sample_id = ceiling(1:1000),
         sample_id = floor(sample_id/20)+1,
         sample_id = ifelse(sample_id == 51, 1, sample_id)) %>% 
  ungroup() 
  

genres_sampled <- genres_sampled %>% 
  mutate(sample_id = paste0(genre, "_", sample_id)) %>% 
  group_by(sample_id) %>% 
  summarise(text = paste0(text_lemm, collapse = "     --     ")) %>% 
  ungroup() %>% 
  mutate(genre = str_remove(sample_id, "_\\d+$"))

str(genres_sampled)
```

```{r}
unique(genres_sampled$genre)
```

Ranks

```{r}
ranks <- genres_sampled %>% 
  unnest_tokens(input = text, output = word, token = "words") %>% 
  count(word, sort = T) %>% 
  head(300)

head(ranks, 10)
tail(ranks, 10)
```

Frequencies (for all genres)

```{r}
genres_freqs <- genres_sampled %>% 
  unnest_tokens(input = text, output = word, token = "words") %>% 
  filter(word %in% ranks$word) %>% 
  group_by(sample_id) %>% 
  count(word) %>% 
  pivot_wider(names_from = word, values_from = n, values_fill = 0) %>% 
  ungroup() 

freqs_scaled <- genres_freqs %>% 
  select(-sample_id) %>% 
  scale() %>% 
  as.tibble()

freqs_scaled <- freqs_scaled %>% 
  mutate(sample_id = genres_freqs$sample_id) 
```

make a loop out of that and test classification

```{r}
f <- freqs_scaled %>% 
  mutate(genre = str_remove(sample_id, "_\\d+$")) %>% 
  select(-sample_id)

table(f$genre)

f1 <- f %>% 
  filter(genre == "no_genre" | genre == "песня")
```

```{r}
glimpse(f1)
```

NB mb classificator would work better with char ngrams

```{r}

```
